For submission to the Journal of Integrated Pest Management

### 2021-08-13 ofv_fife_2021.Rmd
Chaged title, added information about sequence identity

### **Request from Reviewer 4: Revise title to better reflect content of publiction** First report of the *Brevipalpus*-transmitted (Trombidiformes: Tenuipalpidae) orchid fleck virus infecting three ornamentals in Florida

@@

Two orchid-infecting strains of OFV were detected via combinations of conventional RT-PCR, RT-qPCR, Sanger sequencing and High Throughput Sequencing. Amplicons shared 98% nucleotide identity with OFV-Orc1 and OFV-Orc2 available in NCBI GenBank. Coinfections were seen in each county, but single strains of OFV-Orc were seen in *L. muscari* (Alachua, OFV-Orc2) and *A. elatior* (Leon, OFV-Orc1).

@@

These samples from FDACS were subsequently retested by the USDA-ARS, in conjunction with tests of fresh samples from both Alachua and Leon counties. The USDA used RT-PCR, RT-qPCR, and High Throughput Sequencing (HTS) to reconfirm the presence of OFV. RT-PCR and qPCR with Generic R2-Dicho-GF and R2-Dicho-GR primers amplifed ~800 nt of L-gene (RNA2) amplicon [@Roy2020], and OFV-Orc1 and OFV-Orc2 were detected in both *O. intermedius* and *A. elatior* from Leon County. **(insert statement of how much OFV-Orc1 and OFV-Orc2 differ at the nucleotide level)** ***Request from Reviewer 2: deposit and include GenBank accession numbers for the identified virus isolates in Florida? Re:New reports of viruses required that or two other alternative methods for viral detection** **Question from Reviewer 4: Does each isolate reflect a strain?**

@@

The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, *sensu* @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker<!--  (`r citef("fig_4")`) -->.

The first report of OFV in the US is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015].

@@

In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern strains of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the potential threat of different strains of OFV on citrus, which will be a definite concern to the US multi-billion-dollar citrus industry already impacted by the Huanglongbing disease. *B. californicus*, *B. yothersi*, and *B. obovatus* are all present in Florida [@Childers2003; @Akyazi2017] (`r citef("fig_4")`), and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to identify mite populations which need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

### Acknowledgements
We would like to give special thanks to the Tallahassee Museum for their patience, cooperation, and support with collecting plant samples. We also want to thank Drs. Sam Bolton, FDACS and Aline Tassi, Univ. of Sao Paulo, Brazil for checking the mites we have sent for species validation. Furthermore, we are grateful for Dr. Marc S. Frank's identification of the Liriopogons collected. We are especially indebted to the late Dr. Gary Bauchan for his contributions to this study and the field of acarology, he will be greatly missed. This research was partly funded by the USDA National Institute of Food and Agriculture, Hatch project FLA-NFC-005607. Mention of trade names or commercial products in this publication is solely for the purpose of providing specific information and does not imply recommendation or endorsement by the USDA; USDA is an equal opportunity provider and employer.

@@

<!-- `r fig_nums(name = "fig_4", caption = "Florida is home to other common pest species of *Brevipalpus*, which could be confused with *B. californicus*: (a) *B. phoenicis*, dorsal (b) *B. yothersi*, lateral (c) *B. obovatus*, dorsal.")` -->

@@

<!-- ```{r, fig_4, warning=FALSE} -->
<!-- grid::grid.raster(tiff::readTIFF("figure/other_mites.tif")) -->
<!-- ``` -->

>>>>

### Identification of three new ornamental hosts of *Brevipalpus*-transmitted (Trombidiformes: Tenuipalpidae) orchid fleck virus from Florida

@@

Two orchid-infecting strains of OFV were detected via combinations of conventional RT-PCR, RT-qPCR, Sanger sequencing and High Throughput Sequencing. Amplicons shared 98% nucleotide identity with OFV-Orc1 and OFV-Orc2 RNA2 genome sequences available in NCBI GenBank. Coinfections were seen in each county, but single strains of OFV-Orc were seen in *L. muscari* (Alachua, OFV-Orc2) and *A. elatior* (Leon, OFV-Orc1).

@@

These samples from FDACS were subsequently retested by the USDA-APHIS-PPQ S&T Beltsville laboratory, in conjunction with tests of fresh samples from both Alachua and Leon counties. The USDA used RT-PCR, RT-qPCR, and High Throughput Sequencing (HTS) to reconfirm the presence of OFV. RT-PCR and qPCR with Generic R2-Dicho-GF and R2-Dicho-GR primers amplifed ~800 nt of L-gene (RNA2) amplicon [@Roy2020], and OFV-Orc1 and OFV-Orc2 were detected in both *O. intermedius* and *A. elatior* from Leon County. OFV-Orc1 and OFV-Orc2 strains share >90% nt sequence identity, and differ >25% in the L gene nucleotide sequence, indicating two separate species of *Dichoravirus* [@Dietzgen2014].

@@

The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, *sensu* @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker.

The recent report of OFV in the US is thought to be @Ko1985, who describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015].

@@

In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern strains of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the potential threat of different strains of OFV on citrus, which will be a definite concern to the US multi-billion-dollar citrus industry already impacted by the Huanglongbing disease. *B. californicus*, *B. yothersi*, and *B. obovatus* are all present in Florida [@Childers2003; @Akyazi2017], and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to identify mite populations which need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

### Acknowledgements
We would like to give special thanks for the Tallahassee Museum and their patience, cooperation, and support with collecting plant samples. We are grateful for the USDA-APHIS PPQ Beltsville laboratory for their support in the identification and confirmation OFV strains, as well as *Brevipalpus* mite identification. We also want to thank Drs. Sam Bolton, FDACS and Aline Tassi, Univ. of Sao Paulo, Brazil for checking the mites we have sent for species validation. Furthermore, we are grateful for Dr. Marc S. Frank's identification of the Liriopogons collected. We are especially indebted to the late Dr. Gary Bauchan for his contributions to this study and the field of acarology, he will be greatly missed. This research was partly funded by the USDA National Institute of Food and Agriculture, Hatch project FLA-NFC-005607. Mention of trade names or commercial products in this publication is solely for the purpose of providing specific information and does not imply recommendation or endorsement by the USDA; USDA is an equal opportunity provider and employer.

@@

@@
